### 2020-03-08 appendix.tex
misc edits

Given that the number of infected individuals increase exponentially at rate $r$ initially, the equations
for the infectious cases can be rewritten given the ansatz
$E(t)=c_e e^{rt}$,
$I_a(t)=c_a e^{rt}$,
$I_s(t)=c_s e^{rt}$.
Then, it follows that 
\begin{eqnarray}
r c_a&=&p\gamma_e c_e-\gamma_a c_a,\\
r c_s&=&(1-p)\gamma_e c_e-\gamma_s c_s,
\end{eqnarray} 
which implies that
\begin{equation}
\frac{c_a}{c_s}=\frac{p}{1-p}\left[\frac{r+\gamma_s}{r+\gamma_a}\right].
\end{equation}
The prevalence of asymptomatic and symptomatic individuals is different from $p$ and $1-p$ because prevalence measures the individuals that are currently infectious and does not account for individuals that have already recovered.
Finally, the ratio of secondary case production caused by asymptomatic vs.~symptomatic
individuals during the exponential phase should be
\begin{equation}
\frac{q}{1-q} = \left(\frac{\beta_a}{\beta_s}\right)\frac{p}{1-p}\left[\frac{r+\gamma_s}{r+\gamma_a}\right],
\end{equation}
where $q$ is the fraction of new secondary cases caused by
asymptomatic individuals.

>>>>

Given that the number of infected individuals increase exponentially at rate $r$ initially, the equations
for the infectious cases can be rewritten given the ansatz
$E(t)=c_e e^{rt}$,
$I_a(t)=c_a e^{rt}$,
$I_s(t)=c_s e^{rt}$.
Then, it follows that 
\begin{eqnarray}
r c_a&=&p\gamma_e c_e-\gamma_a c_a,\\
r c_s&=&(1-p)\gamma_e c_e-\gamma_s c_s,
\end{eqnarray} 
which implies that
\begin{equation}
\frac{c_a}{c_s}=\frac{p}{1-p}\left[\frac{r+\gamma_s}{r+\gamma_a}\right].
\end{equation}
This shows that the prevalence of asymptomatic and symptomatic individuals is different from $p$ and $1-p$ because prevalence measures the individuals that are currently infectious and does not account for individuals that have already recovered.
Finally, the ratio of secondary case production caused by asymptomatic vs.~symptomatic
individuals during the exponential phase should be
\begin{equation}
\frac{q}{1-q} = \left(\frac{\beta_a}{\beta_s}\right)\frac{p}{1-p}\left[\frac{r+\gamma_s}{r+\gamma_a}\right],
\end{equation}
where $q$ is the fraction of new secondary cases caused by
asymptomatic individuals.

### 2020-03-08 short.body.tex
misc edits

In order to understand how the shape of asymptomatic and symptomatic generation-interval distributions affects $q$ and $\mathcal R_0$, we summarize the distributions in terms of their means and dispersions.
Assuming that the intrinsic generation-interval distributions for asymptomatic and symptomatic cases follow gamma distributions with different means, $\bar G_a$ and $\bar G_s$, and dispersions (represented by squared coefficients of variations), $\kappa_a$ and $\kappa_s$ \citep{park_2019practical}, 
we have
\begin{equation}
\frac{q}{1-q}=\left(\frac{z}{1-z}\right)\left[\frac{(1 + \kappa_s r \bar G_s)^{1/\kappa_s}}{(1 + \kappa_a r \bar G_a)^{1/\kappa_a}}\right].
\label{eq.gammaratio}
\end{equation}
As is apparent from Eq.~\req{eq.qratio}, relevance $q$ increases with $z$, i.e., with both $p$ and $\mathcal R_a/\mathcal R_s$. The consequences for outbreak potential further depends on the relative lengths of generation intervals as well as the relative amount of dispersions in them.
In Supplementary Materials, we use a compartmental ordinary differential equation model to show that the duration of infectious period of asymptomatic individuals affects both $q$ and $\mathcal R_0$, which is a special case of Eq.~\req{eq.qratio}; we also show that prevalence of asymptomatic/symptomatic cases differs from $p$ and $1-p$.

@@

\begin{figure*}[b!]
\begin{center}
\includegraphics[width=0.4\textwidth]{figheatmap.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_R0.pdf}
\caption{Effects of intrinsic proportion of asymptomatic transmission on the relevance of asymptomatic transmission and basic reproduction number, given variation in
the mean generation interval of asymptomatic cases. 
(A) Long/short generation intervals of asymptomatic cases decrease/increase the relevance of asymptomatic cases, $q$.
(B) Long/short generation intervals of asymptomatic cases increase/decrease the basic reproduction number ${\cal{R}}_0$.
Depending on whether $G_a$ is smaller (left side of the white, dashed line) or bigger (right side of the white, dashed line) than $G_s$, naive estimate that does not explicitly account for asymptomatic transmission over- or under-estimates the true $\mathcal R_0$.
The white circles represent the baseline value at $p=0.18$ and $\mathcal R_a = 0.5 \mathcal R_s$.
The white, solid lines represent the contour lines.
Here, we assume $1/r=7$ days, $\bar G_s=8$ days, and $\kappa_s=\kappa_a=0.5$.
\label{fig.relevance}}
\end{center}
\end{figure*}

@@

Figure~\ref{fig.relevance}B compares the effect of relative mean generation interval of asymptomatic cases, $\bar G_a/\bar G_s$, on the basic reproduction number ${\cal{R}}_0$.
When $\bar G_s < \bar G_a$, then ${\cal{R}}_0$ increases with $z$.
This increase arises because asymptomatic infections that have long generation intervals increase the mean of overall generation intervals (see~\citep{park_2019practical}).
Conversely, when $\bar G_s > \bar G_a$ then the strength ${\cal{R}}_0$ decreases with $z$ because short generation intervals of asymptomatic cases reduce the overall mean generation interval.
The qualitative effects of $z$ and $\bar G_a/\bar G_s$ on $q$ and ${\cal{R}}_0$ remain robust when we assume narrower ($\kappa_s = \kappa_a = 0.3$; Figure~S1) or wider ($\kappa_s = \kappa_a = 0.8$; Figure~S2) generation intervals.

As an example, we assume that 18\% of the newly infected cases are asymptomatic \citep{mizumoto_2020} and that the asymptomatic cases are as half as infectious as the symptomatic ones (i.e., $\mathcal R_a = 0.5 \mathcal R_s$).
The intrinsic proportion of asymptomatic transmission $z$ is 0.11 (white circle in Figure~\ref{fig.relevance}A), but asymptomatic cases can account for anywhere between 6\% and 16\% of infection during the exponential growth phase across the range of mean asymptomatic generation interval we explore.
Likewise, the basic reproduction number ranges between 2.3 and 2.6.
Therefore, when ${\cal{R}}_0$ is estimated without explicitly accounting for asymptomatic spread ($\mathcal R_0=2.5$; white circle in Figure~\ref{fig.relevance}B), it can be over- or under- estimated depending on the relative duration of infection between symptomatic and asymptomatic individuals.

% In both cases, the dependence of R0 on q is compared to a baseline of R0 ≈ 3.1 in a model with no secondary case production from asymptomatic infections (see filled circle at bottom left of Figure 1C-D. 
% In other words, when R0 is estimated without explicitly accounting for asymptomatic spread, it can be over- or under- estimated depending on the relative duration of infection between symptomatic and asymptomatic individuals.

>>>>

In order to understand how the shape of asymptomatic and symptomatic generation-interval distributions affects $q$ and $\mathcal R_0$, we summarize the distributions in terms of their means and dispersions.
Assuming that the intrinsic generation-interval distributions for asymptomatic and symptomatic cases follow gamma distributions \citep{park_2019practical} with different means, $\bar G_a$ and $\bar G_s$, and dispersions (represented by squared coefficients of variations), $\kappa_a$ and $\kappa_s$, 
we have
\begin{equation}
\frac{q}{1-q}=\left(\frac{z}{1-z}\right)\left[\frac{(1 + \kappa_s r \bar G_s)^{1/\kappa_s}}{(1 + \kappa_a r \bar G_a)^{1/\kappa_a}}\right].
\label{eq.gammaratio}
\end{equation}
As is apparent from Eq.~\req{eq.qratio}, relevance $q$ increases with $z$, i.e., with both $p$ and $\mathcal R_a/\mathcal R_s$. The consequences for outbreak potential further depends on the relative lengths of generation intervals as well as the relative amount of dispersions in them.
In Supplementary Materials, we use a compartmental ordinary differential equation model to show that the duration of infectious period of asymptomatic individuals affects both $q$ and $\mathcal R_0$, which is a special case of Eq.~\req{eq.qratio}; we also show that prevalence of asymptomatic/symptomatic cases differs from $p$ and $1-p$.

@@

\begin{figure*}[b!]
\begin{center}
\includegraphics[width=0.4\textwidth]{figheatmap.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_R0.pdf}
\caption{Effects of intrinsic proportion of asymptomatic transmission on the relevance of asymptomatic transmission and basic reproduction number, given variation in
the mean generation interval of asymptomatic cases. 
(A) Long/short generation intervals of asymptomatic cases decrease/increase the relevance of asymptomatic cases, $q$.
(B) Long/short generation intervals of asymptomatic cases increase/decrease the basic reproduction number ${\cal{R}}_0$.
Depending on whether $\bar G_a$ is smaller (left side of the white, dashed line) or bigger (right side of the white, dashed line) than $\bar G_s$, naive estimate that does not explicitly account for asymptomatic transmission (white, dashed line) over- or under-estimates the true $\mathcal R_0$.
The white, solid lines represent the contour lines.
Here, we assume $1/r=7$ days, $\bar G_s=8$ days, and $\kappa_s=\kappa_a=0.5$.
\label{fig.relevance}}
\end{center}
\end{figure*}

@@

Figure~\ref{fig.relevance}B compares the effect of relative mean generation interval of asymptomatic cases, $\bar G_a/\bar G_s$, on the basic reproduction number ${\cal{R}}_0$.
When $\bar G_s < \bar G_a$, then ${\cal{R}}_0$ increases with $z$.
This increase arises because asymptomatic infections that have long generation intervals increase the mean of overall generation intervals (see~\citep{park_2019practical}).
Conversely, when $\bar G_s > \bar G_a$ then the strength ${\cal{R}}_0$ decreases with $z$ because short generation intervals of asymptomatic cases reduce the overall mean generation interval.
Therefore, when ${\cal{R}}_0$ is estimated without explicitly accounting for asymptomatic spread (white, dashed line in Figure~\ref{fig.relevance}B), it can be over- or under- estimated depending on the relative duration of infection between symptomatic and asymptomatic individuals.
The qualitative effects of $z$ and $\bar G_a/\bar G_s$ on $q$ and ${\cal{R}}_0$ remain robust when we assume narrower ($\kappa_s = \kappa_a = 0.3$; Figure~S1) or wider ($\kappa_s = \kappa_a = 0.8$; Figure~S2) generation intervals.
